### 2021-11-15 documents/sections/extensions.tex
akuriyo k- review

\subsection{\akuriyo \obj{k-}}
\label{sec:akuriyo}
After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
\textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
He also suggests that \obj{k-} might be more recent, with which we agree; since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
Since there are no examples of or information about \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.

@@

All verbs in \cref{tab:ptirverbs} above did of course not introduce \obj{k-}, either; we do not know the first-person form of the copular verb \obj{eʔi}.
In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
It was not affected by the extension of \akuriyo \obj{k-}, but whether it existed when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).

\ex<tridown> First person forms of \qu{to go down} \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
\begin{tabular}[t]{@{}ll@{}}
\trio & \obj{p-ɨhtə-} \\
\akuriyo & \obj{p-ɨtə-}\\
\end{tabular}
\xe

>>>>

\subsection{\akuriyo \obj{k-}}
\label{sec:akuriyo}
After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
\textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
He also suggests that \obj{k-} might be more recent, with which we agree: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
Since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.

@@

The verbs listed for \PTir in \cref{tab:ptirverbs} above did of course also not introduce \obj{k-} in \akuriyo, although we do not know the first-person form of the copular verb \obj{eʔi}.
In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).

\ex<tridown> First person forms of \qu{to go down}\\
\begin{tabular}[t]{@{}lll@{}}
\trio & \obj{p-ɨhtə-} & \parencites[294]{triomeira1999} \\
\akuriyo & \obj{p-ɨtə-} & \parencite[84]{gildea1994akuriyo}\\
\end{tabular}
\xe
